---
title: "hwsdr functionality"
author: "Koen Hufkens"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    md_extensions: [ 
      "-autolink_bare_uris" 
    ]
vignette: >
  %\VignetteIndexEntry{hwsdr functionality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# load the library
library(ncdf4)
library(raster)
library(sf)
library(hwsdr)

t_sand <- raster::raster(system.file(
  package = "hwsdr",
  "extdata/t_sand.tif"))

# extract values
values <- readRDS(system.file(
  package = "hwsdr",
  "extdata/values.rds"))

```
Programmatic interface to the Harmonized World Soil Database 'HWSD' web services (<https://daac.ornl.gov/cgi-bin/dsviewer.pl?ds_id=1247>). Allows for easy downloads of 'HWSD' soil data directly to your R workspace or your computer. Routines for both single pixel data downloads and gridded data are provided.


## Use

### Single pixel location download

Get world soil values for a single site using the following format, specifying coordinates as a pair of latitude, longitude coordinates. Here the call only extracts the top soil fraction of sand and silt (% weight) for one specific location.

```{r eval = FALSE}
  values <- ws_subset(
    site = "HWSD",
    location = c(34, -81),
    param = c("T_SAND","T_SILT")
  )
```

At this location we have a top soil fraction of sand of 78% weight and a silt fraction of 12 % weight! Data are returned as tidy data frames including basic meta-data of the query for later subsetting.

```{r}
print(values)
```

### Gridded data

You can grab gridded data by specifying a bounding box c(lat, lon, lat, lon) defined by a bottom left and top right coordinates. Here the call only extracts the top soil fraction of sand (%).

```{r eval = FALSE}
  t_sand <- ws_subset(
    site = "HWSD",
    location = c(32, -81, 34, -80),
    param = "T_SAND",
    path = tempdir(),
    internal = TRUE
  )
```

```{r}
raster::plot(t_sand)
```

Alternatively you can use `sf` bounding box (`st_bbox()`) function output to define an extent over which you want to extract gridded data. The structure of the function also allows for pipes to be used.

```{r eval = FALSE}
  a <- sf::st_sf(a = 1:2,
                geom = sf::st_sfc(
                  sf::st_point(c(34, -81)),
                  sf::st_point(c(32, -80))),
                crs = 4326)

  t_sand <- a %>%
    sf::st_bbox() %>%
    ws_subset(param = "T_SAND")
```

This call gives an equivalent dataset as above, as shown in the plot.

```{r}
raster::plot(t_sand)
```
